The Power of Liberty in the American Workplace: Unveiling the Global Impact of Free Will Perceptions on Job Satisfaction

### 2024-02-15 outputs/paper/paper.qmd
finished discussion section

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

@@

## Control Variables Survey Items {#sec-apx-controls}
**Self-Control Scale**
1. I am good at resisting temptation.
2. I have a hard time breaking bad habits. (R)
3. I am lazy. (R)
4. I say inappropriate things. (R)
5. I do certain things that are bad for me, if they are fun. (R)
6. I refuse things that are bad for me.
7. I wish I had more self-discipline. (R)
8. People would say that I have iron self- discipline.
9. Pleasure and fun sometimes keep me from getting work done. (R)
10. I have trouble concentrating.(R)
11. I am able to work effectively toward long-term goals.
12. Sometimes I can’t stop myself from doing something, even if I know it is wrong. (R)
13. I often act without thinking through all the alternatives. (R)

@@

**Locus of Control**
1.
a. Children get into trouble because their patents punish them too much.
b. The trouble with most children nowadays is that their parents are too easy with
them.
7
2.
a. Many of the unhappy things in people's lives are partly due to bad luck.
b. People's misfortunes result from the mistakes they make.
3.
a. One of the major reasons why we have wars is because people don't take enough
interest in politics.
b. There will always be wars, no matter how hard people try to prevent them.
4.
a. In the long run people get the respect they deserve in this world.
b. Unfortunately, an individual's worth often passes unrecognized no matter how hard
he tries.
5.
a. The idea that teachers are unfair to students is nonsense.
b. Most students don't realize the extent to which their grades are influenced by
accidental happenings.
6.
a. Without the right breaks, one cannot be an effective leader.
b. Capable people who fail to become leaders have not taken advantage of their
opportunities.
7.
a. No matter how hard you try, some people just don't like you.
b. People who can't get others to like them don't understand how to get along with
others.
8.
a. Heredity plays the major role in determining one's personality.
b. It is one's experiences in life which determine what they're like.
9.
a. I have often found that what is going to happen will happen.
b. Trusting fate has never turned out as well for me as making a decision to take a
definite course of action.
10.
a. In the case of the well prepared student there is rarely, if ever, such a thing as an
unfair test.
b. Many times, exam questions tend to be so unrelated to course work that studying is
really useless.
11.
a. Becoming a success is a matter of hard work, luck has little or nothing to do with it.
b. Getting a good job depends mainly on being in the right place at the right time.
12.
a. The average citizen can have an influence in government decisions.
b. This world is run by the few people in power, and there is not much the little guy can
do about it.
13.
a. When I make plans, I am almost certain that I can make them work.
b. It is not always wise to plan too far ahead because many things turn out to be a
matter of good or bad fortune anyhow.

@@

**Self-esteem**
1. I feel that I am a person of worth, at least on an equal basis with others.
2. I feel that I have a number of good qualities.
3. All in all, I am inclined to feel that I am a failure. (R)
4. I am able to do things as well as most other people.
5. I feel that I do not have much to be proud of. (R)
6. I take a positive attitude toward myself.
7. On the whole, I am satisfied with myself.
8. I wish I could have more respect for myself. (R)
9. I certainly feel useless at times. (R)
10. At times I think I am no good at all. (R)
Scale: 1 – Strongly disagree; 7 – Strongly agree.

**Self-efficacy**
1. I am strong enough to overcome life's struggles.
2. At root, I am a weak person. (R)
3. I can handle the situations that life brings.
4. I usually feel that I am an unsuccessful person. (R)
5. I often feel that there is nothing that I can do well. (R)
6. I feel competent to deal effectively with the real world.
7. I often feel like a failure. (R)
8. I usually feel I can handle the typical problems that come up in life.
Scale: 1 – Strongly disagree; 7 – Strongly agree.

**Implicit beliefs**
1. The kind of person someone is, is something very basic about them, and it can't be changed
very much.
2. People can do things differently, but the important parts of who they are can't really be
changed.
3. Everyone is a certain kind of person, and there is not much that they can do to really change
that.
4. As much as I hate to admit it, you can't teach an old dog new tricks. People can't really
change their deepest attributes.
5. Everyone, no matter who they are, can significantly change their basic characteristics. (R)
6. People can substantially change the kind of person they are. (R)
7. No matter what kind of person someone is, they can always change very much. (R)
8. People can change even their most basic qualities. (R)
Scale: 1 – Strongly disagree; 7 – Strongly agree.

>>>>

# Discussion {#sec-discussion}

## Effect of Free Will Beliefs
The main result of our analysis is that the belief in free will is sufficient to positively influence one's job satisfaction. Both the Amazon Mechanical Turk and WVS data sources showed significant positive correlations between these two psychological constructs. Free will beliefs are found to be  integral to psychological well-being and motivation, and are commonly linked to agency, autonomy, and responsibility [@wang2009]. By fostering an environment that supports these values, organizations can enhance job satisfaction and performance [@stillman2010]. Moreover, we found that job autonomy is the most important factor in mediating the relationship between free will beliefs and job satisfaction. When employees feel they have the freedom to make decisions and control their work processes, this autonomy reinforces their free will beliefs, leading to higher job satisfaction [@goezukara2016]. Our analysis shows that this applies directly to the workplace, since we found a robust effect of free will beliefs on job satisfaction, even when controlling for other psychological agency constructs, suggests a strong and direct relationship. As @clark2017 have found, belief in free will also alleviates distress and anxiety in regards to punishment and moral judgments, which can further positively impact satisfaction at work. This is crucial information, since it can help us effectively design targeted interventions with long-term benefits for employees and organizations [@feldman2017].

## Demographic Variable Effects
Our findings suggest that while there are slight differences in the median belief in free will among different age groups, these differences do not reach statistical significance. This outcome indicates that the belief in free will might be relatively stable across different stages of adult life, not significantly impacted by age-related factors such as life experience and generational values, but rather by cultural factors @feltz2009. Our findings also suggest that a slight differences based on sex, with women reporting higher beliefs in free will. This finding could reflect underlying gender-related socialization processes, where differing societal expectations and roles for men and women might influence their perceptions of autonomy and control [@baumeister1988]. However, even when found significant, this effect was marginal in both samples, suggesting that the influence of free will beliefs on job satisfaction might operate independently of these demographic factors. This independence has strong implications for potential interventions interventions aimed at enhancing employees' sense of autonomy and control could be broadly applicable and beneficial [@sheldon2001].

Nevertheless, unobserved demographic factors could have also contributed to the significant difference in free will beliefs and job satisfaction ratings between the Amazon MTurk sample and the WVS sample. Latter is much more representative due to its larger sample size, more thorough survey methodology, and more equal distribution of ages and occupations. The Amazon MTurk sample was much more restricted, with a much lower median age, and was not cross-occupational in the same sense the more comprehensive WVS sample is. The nature of Amazon MTurk as a novel and mostly low-wage job platform could also explain the comparatively lower job satisfaction rates and higher beliefs in free will.

## Cultural Differences
Our findings indicate that correlations are now statistically significant for all countries in the sample which suggests a pervasive and possibly strengthening relationship between belief in free will and job satisfaction globally. This contrasts with Feldman et al.'s (2018) findings, where the significance of these correlations may have varied across countries. This change could be indicative of a global trend towards greater recognition of the importance of autonomy and personal agency in job satisfaction [@goezukara2016]. Our finding that the United States has a high average belief in free will, second only to Mexico, can be contextualized within the framework of cultural values and individualism. Based on how American cultural norms and values are shaped by freedom, independence, and liberty, the perception and valuation of free will can be expected to be considerably higher than when compared to other countries, influencing the strength of its relationship with job satisfaction. Hofstede's cultural dimensions theory, which explores the impact of a society's culture on the values of its members and how these values relate to behavior, suggests that societies with high individualism scores tend to place a greater emphasis on personal freedom and autonomy, which could contribute to stronger beliefs in free will [@hofstede2001]. Again, this has direct implications for interventions operating within the American context, suggesting a strong focus on such values and constructs to increase overall worker's satisfaction.

## Limitations
Potential weaknesses include the overreliance on self-reporting measures, which are subject to biases; the cross-sectional nature of some studies, which limits causal inference; and the potential for cultural nuances to be oversimplified or to not be fully captured by the statistical measures used.  Additionally, the concept of free will can vary significantly across different cultures. When dealing with constructs that deeply intertwine with personal identity and cultural beliefs it is very important to design survey questions appropriately and make sure to not overstep any boundaries. However, the resulting general nature of these questions could limit capturing the full scope of the concept. The standardized scales used to measure the constructs may have therefore not adequately captured the underlying construct, despite high reliability coefficients.

## Future Directions
The next steps to take would be to further explore societal changes over time and individual differences in the value of free will, which would be possible due to the longitudinal nature of the WVS data. Further research couldalso  investigate the impact of emerging technologies and digital platforms on perceptions of free will, examining how online behavior and the emergence of artificial intelligence influences or reflects these beliefs. Additionally, interdisciplinary approaches combining psychology, sociology, and neuroscience could offer deeper insights into the biological underpinnings and social determinants of free will beliefs. Exploring these areas could significantly enhance our understanding of free will's role in modern society and individual psychology, and help us even more in designing appropriate interventions not only at the workspace but other spheres of our lives.

@@

## Control Variables Survey Items {#sec-apx-controls}
**Self-Control Scale**

1. I am good at resisting temptation.
2. I have a hard time breaking bad habits. (R)
3. I am lazy. (R)
4. I say inappropriate things. (R)
5. I do certain things that are bad for me, if they are fun. (R)
6. I refuse things that are bad for me.
7. I wish I had more self-discipline. (R)
8. People would say that I have iron self- discipline.
9. Pleasure and fun sometimes keep me from getting work done. (R)
10. I have trouble concentrating.(R)
11. I am able to work effectively toward long-term goals.
12. Sometimes I can’t stop myself from doing something, even if I know it is wrong. (R)
13. I often act without thinking through all the alternatives. (R)

@@

**Locus of Control**

1.
a. Children get into trouble because their patents punish them too much.
b. The trouble with most children nowadays is that their parents are too easy with
them.
7
2.
a. Many of the unhappy things in people's lives are partly due to bad luck.
b. People's misfortunes result from the mistakes they make.
3.
a. One of the major reasons why we have wars is because people don't take enough
interest in politics.
b. There will always be wars, no matter how hard people try to prevent them.
4.
a. In the long run people get the respect they deserve in this world.
b. Unfortunately, an individual's worth often passes unrecognized no matter how hard
he tries.
5.
a. The idea that teachers are unfair to students is nonsense.
b. Most students don't realize the extent to which their grades are influenced by
accidental happenings.
6.
a. Without the right breaks, one cannot be an effective leader.
b. Capable people who fail to become leaders have not taken advantage of their
opportunities.
7.
a. No matter how hard you try, some people just don't like you.
b. People who can't get others to like them don't understand how to get along with
others.
8.
a. Heredity plays the major role in determining one's personality.
b. It is one's experiences in life which determine what they're like.
9.
a. I have often found that what is going to happen will happen.
b. Trusting fate has never turned out as well for me as making a decision to take a
definite course of action.
10.
a. In the case of the well prepared student there is rarely, if ever, such a thing as an
unfair test.
b. Many times, exam questions tend to be so unrelated to course work that studying is
really useless.
11.
a. Becoming a success is a matter of hard work, luck has little or nothing to do with it.
b. Getting a good job depends mainly on being in the right place at the right time.
12.
a. The average citizen can have an influence in government decisions.
b. This world is run by the few people in power, and there is not much the little guy can
do about it.
13.
a. When I make plans, I am almost certain that I can make them work.
b. It is not always wise to plan too far ahead because many things turn out to be a
matter of good or bad fortune anyhow.

@@

**Self-esteem**

1. I feel that I am a person of worth, at least on an equal basis with others.
2. I feel that I have a number of good qualities.
3. All in all, I am inclined to feel that I am a failure. (R)
4. I am able to do things as well as most other people.
5. I feel that I do not have much to be proud of. (R)
6. I take a positive attitude toward myself.
7. On the whole, I am satisfied with myself.
8. I wish I could have more respect for myself. (R)
9. I certainly feel useless at times. (R)
10. At times I think I am no good at all. (R)
Scale: 1 – Strongly disagree; 7 – Strongly agree.

**Self-efficacy**

1. I am strong enough to overcome life's struggles.
2. At root, I am a weak person. (R)
3. I can handle the situations that life brings.
4. I usually feel that I am an unsuccessful person. (R)
5. I often feel that there is nothing that I can do well. (R)
6. I feel competent to deal effectively with the real world.
7. I often feel like a failure. (R)
8. I usually feel I can handle the typical problems that come up in life.
Scale: 1 – Strongly disagree; 7 – Strongly agree.

**Implicit beliefs**

1. The kind of person someone is, is something very basic about them, and it can't be changed
very much.
2. People can do things differently, but the important parts of who they are can't really be
changed.
3. Everyone is a certain kind of person, and there is not much that they can do to really change
that.
4. As much as I hate to admit it, you can't teach an old dog new tricks. People can't really
change their deepest attributes.
5. Everyone, no matter who they are, can significantly change their basic characteristics. (R)
6. People can substantially change the kind of person they are. (R)
7. No matter what kind of person someone is, they can always change very much. (R)
8. People can change even their most basic qualities. (R)
Scale: 1 – Strongly disagree; 7 – Strongly agree.
